How to Analyze Build Logs

### 2021-04-19 tse-paper/paper.tex
simplify technique descriptions

To enable configurable build log analysis, we evaluate the following
techniques:
\begin{itemize}
  \item \textbf{\emph{Programming by Example (PBE)}}, which learns
  regular expressions based on few user-given log excerpts.
  We rely on the PROSE Framework \caro{link!} that is also responsible for auto-filling
  cells in Microsoft Excel.

@@

\subsection{Programming by Example (PBE)}
The idea behind \emph{Programming by Example} is to automatically
generate a program, by capturing the
intent of the user through examples that the user provides.
Leveraging this approach, we designed a chunk retrieval technique
for build logs.
The \emph{PROSE library}\footnote{https://microsoft.github.io/prose/}
forms the basis for our implementation.
This library builds on the generic program synthesis framework
\emph{FlashMeta}~\citep{polozov2015flashmeta:} and the specialized
text extraction DSL \emph{FlashExtract}~\citep{le2014flashextract:}.
% Both enable \emph{PROSE} to synthesize regular expression programs
% consistent~\citep{mitchell1982generalization} with a set of in-
% and output examples.
% The usage of example enables the user to configure a chunk
% retrieval without understanding the whole structure of
% the build log.

\subsubsection{Configuration}
In- and output example pairs are the main driver of PBE;
The \emph{input} is the full text of the build log file.
The \emph{output} is
a substring of the log file text, representing the
substring that the synthesized program should retrieve when
given the corresponding input file.
One or multiple examples, the
training set, \emph{configure} a specific chunk retrieval with PBE:
they define the chunk of a build log that should be retrieved.
The PROSE program synthesis then tries to construct a program
consistent with all training
examples~\citep{mitchell1982generalization}.
If PROSE cannot synthesize a program, e.g.,\
because the required regular expression
is too complicated to be synthesized, PBE returns the
error message ``no program found.''

\subsubsection{Chunk Retrieval Run}
A run of PBE takes a build log file as input and executes the
synthesized regular expression program on it.
It then returns the substring
of the build log matched by the program.
If the program finds no match because the analyzed build log
does not contain the structure that the training examples define,
PBE returns the error message ``no match found.''

\subsection{Common Text Similarity (CTS)}
Text Similarity approaches are widely used to filter unstructured
textual software artifacts~\citep{runeson2007detection,
marcus2005recovery,antoniol2002recovering,mccarey2006recommending}.
These approaches fall into the broader area of information retrieval
techniques, which several of the works from our literature mapping
used to analyze build logs.
Similar to these approaches, we designed a chunk retrieval
technique that
retrieves those lines from a build log which are most
similar to the specified examples.

\subsubsection{Configuration}
To configure CTS we chose to use the same concept of examples as for
PBE: lines of the targeted chunks from the training examples define
the search query.
Using a Vector Space
Model~\citep{schutze2008introduction} (VSM), an algebraic model for
representing textual documents, we encode each line as a separate term
vector.
The terms that make up the dictionary are simply all words
that appear in the example logs.
We separate words by spaces and
punctuation.
The vectors comprise points counting
how many times a word out of the dictionary appears in the specific
line at hand.

@@

\subsubsection{Chunk Retrieval Run}
To retrieve the desired information from a build log, we encode every
line of the input build log in VSM.
The algorithm calculates the cosine
similarity to compare each line of the
build log with each line of the search query.
After summing up the
similarities of each build log line to all search query lines, we sort
the build log lines in decreasing similarity.
The average number of
lines in the chunks of the training examples determines how many of
the most similar lines our approach returns as the output of the retrieval
run.

@@

\subsubsection{Configuration}
A set of keywords configures the chunk retrieval with KWS\@.
To better
compare KWS with PBE and CTS, we also configure it through examples.
From each example, we extract a list of keywords that appeared in the
training examples and that we expect to appear
in, or close to, the targeted chunk in a new build log.
Finally, we merge the lists from all examples to build one unified
list of the most commonly occurring keywords for the search with KWS.

\subsubsection{Chunk Retrieval Run}
For a retrieval run, we take an entire build log file as input and
search for all exact occurrences of the search keywords.
As keywords are
often not directly describing the desired information, but rather
appear close to it, KWS also retrieves a context of lines
around the found keyword.
The number of surrounding lines retrieved is
the average number of lines in the chunks of the training examples.

\subsection{Random Line Retrieval (RLR)}
\label{sec:expl-rlr}
To fully comprehend how difficult the task of build log analysis really
is without any a priori knowledge, we include a comparison with a
baseline of
picking lines at random from the build log.
RLR mimics the
situation of guessing blindly which lines are interesting.
The only configuration option for RLR is the number of lines it should
retrieve.
For a fair comparison to the other techniques (whose number of
returned lines is largely determined by the training set),
we configure RLR to return the average
number of lines in the chunks of the training examples.

>>>>

To enable configurable build log analysis, we evaluate the following
techniques:
\begin{itemize}
  \item \textbf{\emph{Programming by Example (PBE)}}, which learns
  regular expressions based on few user-given log excerpts.
  We rely on the PROSE
  Framework\footnote{https://microsoft.github.io/prose/},
  which also provides the auto-filling for
  cells in Microsoft Excel.

@@

\todo{Intro section? In this section we discuss the details of our
prototypes that apply the three chosen techniques to build log
analysis.}

\subsection{Programming by Example (PBE)}
The idea behind \emph{Programming by Example} is to automatically
generate a program, by capturing the
intent of the user through examples that the user provides.
Leveraging this approach, we designed a chunk retrieval technique
for build logs.
The \emph{PROSE library}\footnote{https://microsoft.github.io/prose/}
forms the basis for our implementation.
This library builds on the generic program synthesis framework
\emph{FlashMeta}~\citep{polozov2015flashmeta:} and the specialized
text extraction DSL \emph{FlashExtract}~\citep{le2014flashextract:}.

To configure our prototype, the user supplies one or more build logs
together with the excerpt of these logs they would like to be
retrieved.
The PROSE program synthesis then tries to construct a program
consistent with all training
examples~\citep{mitchell1982generalization}.
If PROSE cannot synthesize a program, e.g.,\
because the required regular expression
is too complicated to be synthesized, our prototype returns the
error message ``no program found.''

When our PBE prototype analyzes a build log, it tries to match
the synthesized regular expression program.
It then returns the matched substring.
If there is no match, because the analyzed build log
does not contain the structure that the training examples define,
PBE returns the error message ``no match found.''

\subsection{Common Text Similarity (CTS)}
Text Similarity approaches are widely used to filter unstructured
textual software artifacts~\citep{runeson2007detection,
marcus2005recovery,antoniol2002recovering,mccarey2006recommending}.
These approaches fall into the broader area of information retrieval
techniques, which several of the works from our literature mapping
used to analyze build logs.
Similar to these approaches, we designed a chunk retrieval
technique that
retrieves those lines from a build log which have a similar
vocabulary to the log excerpts the user provides as configuration.

Using a Vector Space
Model~\citep{schutze2008introduction} (VSM), an algebraic model for
representing textual documents, we encode each line of the
user-provided log excerpts as a separate term vector.
The terms that make up the dictionary are simply all words
that appear in the example logs.
We separate words by spaces and
punctuation.
The vectors comprise points counting
how many times a word out of the dictionary appears in the specific
line at hand.

@@

To analyze a build log, we encode every
line of the build log in VSM.
The algorithm calculates the cosine
similarity to compare each line of the
build log with each line of the search query.
After summing up the
similarities of each build log line to all search query lines, we sort
the build log lines in decreasing similarity.
The average number of
lines in the chunks of the training examples determines how many of
the most similar lines our approach returns as the output of the retrieval
run.

@@

The user provides a list of search terms to configure KWS.
To analyze a build log, our prototype searches for all
exact occurrences of the search keywords.

To better
compare KWS with the other two techniques,
we also configure it through a set of log excerpts.
From each excerpts, we manually extract a list of keywords that
appear in the
training examples and that we expect to appear
in, or close to, the targeted chunk in a new build log.
Finally, we merge the lists from all examples to build one unified
list of the most commonly occurring keywords for the search with KWS.
As keywords are
often not directly describing the desired information, but rather
appear close to it, KWS also retrieves a context of lines
around the found keyword.
The number of surrounding lines retrieved is
the average number of lines in the chunks of the training examples.

\subsection{Random Line Retrieval (RLR)}
\label{sec:expl-rlr}
To fully comprehend how difficult the task of build log analysis really
is without any a priori knowledge, we include a comparison with a
baseline of
picking lines at random from the build log.
RLR mimics the
situation of guessing blindly which lines are interesting.
The only configuration option for RLR is the number of lines it should
retrieve.
For a fair comparison to the other techniques (whose number of
returned lines is largely determined by the training set),
we let RLR return the average
number of lines in the log excerpts used as configuration.
